Evolution of precopulatory and postcopulatory inbreeding behaviour

### 2016-09-05 ms_pre_post.Rmd
Start introductory cost paragraph, and separate polyandry and inbreeding mating strategy costs in the Model

<!--- TODO: PARAGRAGH ON IDEA THAT RELATIVE COSTS WILL DETERMINE PRE VS POST --->
Evolution of inbreeding fertilisation stratey might increase selection for polyandry, but evolutionary feedbacks between inbreeding fertilisation strategy and polyandry might also be contingent upon evolution of inbreeding mating strategy, thereby leading to complex evolutionary dynamics between inbreeding strategies and polyandry. For example, if females can acheive their desired degree of inbreeding through mate choice alone, then subsequent polyandry and inbreeding fertilisation strategy might be superfluous for increasing female fitness. Alternatively, if females are often unsuccessful at realising their inbreeding mating strategy due to factors such as sexual conflict over mating and resulting male coercion [e.g., @Parker2006], limited kin recognition amongst potential mates [e.g., @Penn2010], or simply the inherent stochasticity associated with mate encounters or expression of mate choice based on inbreeding [e.g., @Kokko2006; @Duthie], then selection on inbreeding fertilisation strategy might occur to reinforce inbreeding mating strategy. In empirical systems where inbreeding depression is severe and inbreeding avoidance is therefore beneficial, expression of inbreeding mating strategy and inbreeding fertilisation strategy have been observed across diverse taxa [e.g., @Gasparini2011; @Tuni2013; @Firman2015; @Fischer2015; @Daniel2015]. However, no models predict *a priori* which of these inbreeding strategies should be observed when inbreeding depression is sufficiently severe to make inbreeding avoidance adaptive.

<!--- TODO: PARAGRAGH ON COSTS OF POLYANDRY, MATING, AND FERTILISATION STRATEGIES --->

<!--- TODO: OVERVIEW OF THE POINT OF THE MODELLING --->
Here we apply a model of female and male reproductive interactions and associated inbreeding depression in offspring viability to quantify evolution of three separate traits including female tendency for polyandry, female pre-copulatory inbreeding strategy (mating strategy), and female post-copulatory inbreeding strategy (fertilisation strategy). We use *in silico* experiments to test the hypotheses that, given inbreeding depression in offspring viability, evolution of post-copulatory inbreeding avoidance causes selection for evolution of polyandry, and that pre-copulatory but not post-copulatory inbreeding avoidance will evolve when pre-copulatory inbreeding avoidance is less costly than the combined cost of post-copulatory inbreeding avoidance and polyandry, while post-copulatory but not pre-copulatory inbreeding avoidance will evolve when pre-copulatory inbreeding avoidance is more costly than the combined costs of post-copulatory inbreeding avoidance and polyandry. We thereby provide a framework for predicting evolution of pre-copulatory and post-copulatory inbreeding avoidance across reproductive systems with different relative costs and different magnitudes of inbreeding depression in offspring viability.

@@

Values of phenotypic traits for tendency to engage in polyandry ($P_{p}$), mating ($M_{p}$), and fertilistation ($F_{p}$) each have costs, which independently increase the probability of a focal female's reproductive failure. These costs are all expressed prior to mating. In the case of tendency to engage in polyandry and inbreeding mating strategy, costs of $P_{p}$ ($c_{P}$) and $M_{p}$ ($c_{M}$) are interpreted biologically as costs associated with mate searching or courtship [e.g., @Rowe1988; @Rowe1994; @Ronkainen1994; @Koga1998; @Lasley-Rasher2012]. In the case of inbreeding fertilisation strategy, the cost of $F_{p}$ ($c_{F}$) is interpreted as an energetic cost associated with the physiological or biochemical mechanisms that need to be developed to bias fertilsation for inbreeding avoidance or inbreeding preference [e.g., @Gasparini2011]. Each cost independently increases the risk of female pre-mating mortality and, because generations are non-overlapping, total reproductive failure as a linear function of its associated trait. The probability that a focal female dies before mating due to the cost of polyandry is therefore $P_{p} \times c_{P}$. Similarly, the probability of pre-mating female mortality due to the cost of mating strategy and fertilisation strategy is $|M_{p}| \times c_{M}$ and $|F_{p}| \times c_{F}$, repectively, where $|M_{p}|$ is the absolute value of $M_{p}$ and $|F_{p}|$ is the absolute value of $F_{p}$. Absolute values are used for applying costs to inbreeding strategies because both negative and positive $M_{p}$ and $F_{p}$ bias breeding interactions with respect to kinship, while only positive values of $P_{p}$ affect a female's tendency for polyandry; zero and negative $P_{p}$ values always translate to monandrous phenotypes (see below).

>>>>

<!--- TODO: PARAGRAGH ON IDEA OF PRE VS POST COPULATORY INBREEDING STRATEGY --->
Evolution of inbreeding fertilisation stratey might increase selection for polyandry, but evolutionary feedbacks between inbreeding fertilisation strategy and polyandry might also be contingent upon evolution of inbreeding mating strategy, thereby leading to complex evolutionary dynamics between inbreeding strategies and polyandry. For example, if females can acheive their desired degree of inbreeding through mate choice alone, then subsequent polyandry and inbreeding fertilisation strategy might be superfluous for increasing female fitness. Alternatively, if females are often unsuccessful at realising their inbreeding mating strategy due to factors such as sexual conflict over mating and resulting male coercion [e.g., @Parker2006], limited kin recognition amongst potential mates [e.g., @Penn2010], or simply the inherent stochasticity associated with mate encounters or expression of mate choice based on inbreeding [e.g., @Kokko2006; @Duthie], then selection on inbreeding fertilisation strategy might occur to reinforce inbreeding mating strategy. In empirical systems where inbreeding depression is severe and inbreeding avoidance is therefore beneficial, expression of inbreeding mating strategy and inbreeding fertilisation strategy have been observed across diverse taxa [e.g., @Gasparini2011; @Tuni2013; @Firman2015; @Fischer2015; @Daniel2015]. However, no models predict *a priori* which of these inbreeding strategies should be observed when inbreeding depression is sufficiently severe to make inbreeding avoidance adaptive.

<!--- TODO: PARAGRAGH ON COSTS OF POLYANDRY, MATING, AND FERTILISATION STRATEGIES --->
Evolution of polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy will ultimately be determined by the, potentially interacting, fitness benefits and costs of each phenotype. Given severe inbreeding depression, phenotypes will increase a focal female's fitness by causing her to avoid breeding with close relatives. However, evolution of phenotypes will also be directly affected by negative selection (i.e., costs) associated with the tendency to engage in polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy.

<!--- Mate searching: Lehmann2013, Kokko2006, Jamieson2009. Kin discrimination: Waser1986. Increased predation risk associated with mate search: Rowe1988, Ronkainen1994, Koga1998. Increased risk associated with expression of mate choice (e.g., avoiding male coercion): Rowe1994. Risk of not mating due to choosiness: Kokko2013

<!--- TODO: OVERVIEW OF THE POINT OF THE MODELLING --->
Here we apply a model of female and male reproductive interactions and associated inbreeding depression in offspring viability to quantify evolution of three separate traits including female tendency for polyandry, female pre-copulatory inbreeding strategy (mating strategy), and female post-copulatory inbreeding strategy (fertilisation strategy). We use individual-based modelling and *in silico* experiments to test the hypotheses that, given inbreeding depression in offspring viability, evolution of post-copulatory inbreeding avoidance causes selection for evolution of polyandry, and that pre-copulatory but not post-copulatory inbreeding avoidance will evolve when pre-copulatory inbreeding avoidance is less costly than the combined cost of post-copulatory inbreeding avoidance and polyandry, while post-copulatory but not pre-copulatory inbreeding avoidance will evolve when pre-copulatory inbreeding avoidance is more costly than the combined costs of post-copulatory inbreeding avoidance and polyandry. We thereby provide a framework for predicting evolution of pre-copulatory and post-copulatory inbreeding avoidance across reproductive systems with different relative costs and different magnitudes of inbreeding depression in offspring viability.

@@

<!--- Change mating cost to an increased risk of mating failure ---> 
Values of phenotypic traits for tendency to engage in polyandry ($P_{p}$), mating ($M_{p}$), and fertilistation ($F_{p}$) each have costs, which independently increase the probability of a focal female's reproductive failure. These costs are all expressed prior to mating. In the case of tendency to engage in polyandry, the cost of $P_{p}$ ($c_{P}$) is interpreted biologically as a cost associated with mate searching or courtship [e.g., @Rowe1988; @Rowe1994; @Ronkainen1994; @Koga1998; @Lasley-Rasher2012]. For inbreeding mating strategy, the cost of $M_{p}$ ($c_{M}$) is interpreted biologically as a cost associated with risk of mating failure due to females being overly choosy [e.g., @Kokko2013], and for inbreeding fertilisation strategy, the cost of $F_{p}$ ($c_{F}$) is interpreted as an energetic cost associated with the physiological or biochemical mechanisms that need to be developed to bias fertilsation for inbreeding avoidance or inbreeding preference [e.g., @Gasparini2011]. Each cost independently increases the risk of female pre-mating mortality and, because generations are non-overlapping, total reproductive failure as a linear function of its associated trait. The probability that a focal female dies before mating due to the cost of polyandry is therefore $P_{p} \times c_{P}$. Similarly, the probability of pre-mating female mortality due to the cost of mating strategy and fertilisation strategy is $|M_{p}| \times c_{M}$ and $|F_{p}| \times c_{F}$, repectively, where $|M_{p}|$ is the absolute value of $M_{p}$ and $|F_{p}|$ is the absolute value of $F_{p}$. Absolute values are used for applying costs to inbreeding strategies because both negative and positive $M_{p}$ and $F_{p}$ bias breeding interactions with respect to kinship, while only positive values of $P_{p}$ affect a female's tendency for polyandry; zero and negative $P_{p}$ values always translate to monandrous phenotypes (see below).
